---
title: "Gene-Level Visualization"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Gene-Level Visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 10,
  fig.height = 6,
  out.width = "100%"
)
```

## Overview

This article covers `PlotGene()` and `PlotRegion()` in depth — the two
functions for producing peak-on-gene-structure diagrams.

```{r load}
library(RNAPeaks)
data(gtf_human)
```

---

## PlotGene()

### Basic usage

```{r plot-gene-basic, eval = FALSE}
result <- PlotGene(
    bed    = sample_bed,
    geneID = "GAPDH",
    gtf    = gtf_human
)
result$plot
```

### Selecting a transcript isoform

By default, RNAPeaks selects the longest transcript. To target a specific
isoform, supply its Ensembl transcript ID:

```{r plot-gene-isoform, eval = FALSE}
result <- PlotGene(
    bed    = sample_bed,
    geneID = "TP53",
    gtf    = gtf_human,
    TxID   = "ENST00000269305"
)
result$plot
```

### Ordering protein tracks

The `order_by` argument controls the vertical stacking order of RBP tracks:

```{r order-by, eval = FALSE}
# Order by peak count (most peaks at the bottom — default)
PlotGene(bed = sample_bed, geneID = "GAPDH", gtf = gtf_human, order_by = "Count")

# Alphabetical order
PlotGene(bed = sample_bed, geneID = "GAPDH", gtf = gtf_human, order_by = "Target")

# Explicit custom order
PlotGene(
    bed      = sample_bed,
    geneID   = "GAPDH",
    gtf      = gtf_human,
    order_in = c("PROTEIN_A", "PROTEIN_B", "PROTEIN_C")
)
```

### Merging nearby peaks

Use `merge` to collapse peaks within a given distance (bp) into a single
rectangle, reducing visual noise from fragmented binding sites:

```{r merge, eval = FALSE}
PlotGene(
    bed    = sample_bed,
    geneID = "GAPDH",
    gtf    = gtf_human,
    merge  = 50     # merge peaks within 50 bp
)
```

### 5′ → 3′ orientation

For negative-strand genes, the x-axis runs right-to-left by default
(genomic coordinate order). Set `five_to_three = TRUE` to flip the axis
so 5′ is always on the left:

```{r five-to-three, eval = FALSE}
PlotGene(
    bed           = sample_bed,
    geneID        = "ACTB",
    gtf           = gtf_human,
    five_to_three = TRUE
)
```

### Highlighting a region

Draw a shaded rectangle over any genomic interval to call attention to a
region of interest:

```{r highlight, eval = FALSE}
PlotGene(
    bed                       = sample_bed,
    geneID                    = "GAPDH",
    gtf                       = gtf_human,
    highlighted_region_start  = 6643000,
    highlighted_region_stop   = 6643500,
    highlighted_region_color  = "gold",
    highlighted_region_opacity = 0.4
)
```

### Exon-intron junction lines

Add vertical dashed lines at every exon-intron boundary:

```{r junctions, eval = FALSE}
PlotGene(
    bed            = sample_bed,
    geneID         = "GAPDH",
    gtf            = gtf_human,
    show_junctions = TRUE,
    junction_color = "steelblue"
)
```

---

## PlotRegion()

`PlotRegion()` renders all genes (using their longest transcript) whose
features overlap the specified genomic window, stacked vertically.

### Basic usage

```{r plot-region, eval = FALSE}
result <- PlotRegion(
    bed    = sample_bed,
    gtf    = gtf_human,
    Chr    = "12",
    Start  = 56000000,
    End    = 56050000,
    Strand = "+"
)
result$plot
```

### Focusing on a single gene within a region

Supply `geneID` to include only one gene while still using the specified
coordinate window for peak filtering:

```{r region-gene, eval = FALSE}
PlotRegion(
    bed    = sample_bed,
    gtf    = gtf_human,
    Chr    = "12",
    Start  = 56000000,
    End    = 56050000,
    Strand = "+",
    geneID = "GAPDH"
)
```

---

## Styling reference

Both `PlotGene()` and `PlotRegion()` accept a shared set of styling
arguments via `...`. The most commonly used are listed below.

### Gene structure

| Parameter | Default | Description |
|-----------|---------|-------------|
| `exon_fill` | `"navy"` | Exon / CDS fill color |
| `utr_fill` | `"lightgray"` | UTR fill color |
| `intron_color` | `"gray60"` | Intron line color |
| `intron_linewidth` | `0.9` | Intron line width |

### Peaks

| Parameter | Default | Description |
|-----------|---------|-------------|
| `peak_col` | `"purple"` | Peak rectangle fill color |
| `peak_alpha` | `0.95` | Peak opacity |
| `peak_border_color` | `NA` | Peak border color (`NA` = none) |
| `peaks_width` | `0.3` | Vertical height of each peak row |

### Labels

| Parameter | Default | Description |
|-----------|---------|-------------|
| `label_size` | `5` | Protein label font size |
| `label_color` | `"black"` | Protein label color |
| `title_size` | `25` | Plot title font size |

### Layout

| Parameter | Default | Description |
|-----------|---------|-------------|
| `axis_pad_bp` | `500` | Padding (bp) added to each side of the plot |
| `axis_breaks_n` | `5` | Number of x-axis tick marks |
| `max_proteins` | `40` | Maximum number of protein tracks to show |
| `x_lims` | `NULL` | Custom x-axis limits as `c(min, max)` |

### Background bands

| Parameter | Default | Description |
|-----------|---------|-------------|
| `band_even_fill` | `"#F7F8FA"` | Even-row band fill |
| `band_odd_fill` | `"#FFFFFF"` | Odd-row band fill |
| `band_sep_color` | `"#E5E7EB"` | Band separator line color |

---

## Saving output

Both functions write a PDF and a CSV by default. Set paths explicitly or
pass `NULL` to suppress file output:

```{r save, eval = FALSE}
result <- PlotGene(
    bed                = sample_bed,
    geneID             = "GAPDH",
    gtf                = gtf_human,
    RNA_Peaks_File_Path = "figures/GAPDH_peaks.pdf",
    Bed_File_Path       = "data/GAPDH_peaks.csv"
)

# Suppress file output — work with the ggplot object directly
result <- PlotGene(
    bed                = sample_bed,
    geneID             = "GAPDH",
    gtf                = gtf_human,
    RNA_Peaks_File_Path = NULL,
    Bed_File_Path       = NULL
)

# Save manually with custom dimensions
ggplot2::ggsave("GAPDH.pdf", result$plot, width = 18, height = 10)
```

---

## Session info

```{r session-info}
sessionInfo()
```
