---
title: "PlotRegion"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{PlotRegion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 10,
  fig.height = 6,
  out.width = "100%"
)
```

## Overview

`PlotRegion()` renders all genes whose features overlap a specified genomic
window, stacking each gene's transcript structure and overlapping RBP peaks
vertically. It is useful for visualizing binding patterns across a locus or
comparing multiple genes in a region simultaneously.

```{r load}
library(RNAPeaks)
data(gtf_human)
```

---

## Basic usage

```{r basic, eval = FALSE}
result <- PlotRegion(
    bed    = sample_bed,
    gtf    = gtf_human,
    Chr    = "12",
    Start  = 56000000,
    End    = 56050000,
    Strand = "+"
)
result$plot
```

---

## Focusing on a single gene within a region

Supply `geneID` to include only one gene while still using the coordinate
window for peak filtering:

```{r single-gene, eval = FALSE}
PlotRegion(
    bed    = sample_bed,
    gtf    = gtf_human,
    Chr    = "12",
    Start  = 56000000,
    End    = 56050000,
    Strand = "+",
    geneID = "GAPDH"
)
```

---

## Saving output

```{r save, eval = FALSE}
result <- PlotRegion(
    bed                 = sample_bed,
    gtf                 = gtf_human,
    Chr                 = "12",
    Start               = 56000000,
    End                 = 56050000,
    Strand              = "+",
    RNA_Peaks_File_Path = "figures/region_peaks.pdf",
    Bed_File_Path       = "data/region_peaks.csv"
)

# Suppress file output
PlotRegion(
    bed                 = sample_bed,
    gtf                 = gtf_human,
    Chr                 = "12",
    Start               = 56000000,
    End                 = 56050000,
    Strand              = "+",
    RNA_Peaks_File_Path = NULL,
    Bed_File_Path       = NULL
)
```

