---
title: "createRetainedIntronSplicingMap"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{createRetainedIntronSplicingMap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 10,
  fig.height = 5,
  out.width = "100%"
)
```

## Overview

`createRetainedIntronSplicingMap()` quantifies RBP binding frequency around
the two junctions of retained-intron (RI) events: the upstream exon/intron
boundary and the downstream intron/exon boundary. It accepts the same
SE.MATS column format and the same parameters as `createSplicingMap()`.

```{r load}
library(RNAPeaks)
```

---

## Basic usage

```{r basic, eval = FALSE}
createRetainedIntronSplicingMap(
    bed_file = sample_bed,
    RIMATS   = sample_se.mats
)
```

---

## Event classification

Events are classified using the same thresholds as `createSplicingMap()`:

- **Retained** — `IncLevelDifference < -0.1`
- **Excluded** — `IncLevelDifference > 0.1`
- **Control** — non-significant, stable inclusion

```{r thresholds, eval = FALSE}
createRetainedIntronSplicingMap(
    bed_file                     = sample_bed,
    RIMATS                       = sample_se.mats,
    retained_IncLevelDifference  = -0.1,
    exclusion_IncLevelDifference = 0.1,
    p_valueRetainedAndExclusion  = 0.05,
    p_valueControls              = 0.95,
    Min_Count                    = 50
)
```

---

## Bootstrap control and significance

The same bootstrap z-score approach is used as in `createSplicingMap()`.
See the [createSplicingMap article](splicing-maps.html) for full details.

```{r significance, eval = FALSE}
createRetainedIntronSplicingMap(
    bed_file           = sample_bed,
    RIMATS             = sample_se.mats,
    control_iterations = 30,
    use_fdr            = TRUE,
    fdr_threshold      = 0.05,
    show_significance  = TRUE
)
```

---

## Returning data

```{r return-data, eval = FALSE}
freq_df <- createRetainedIntronSplicingMap(
    bed_file    = sample_bed,
    RIMATS      = sample_se.mats,
    return_data = TRUE
)
head(freq_df)
```

