---
title: "createRetainedIntronSequenceMap"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{createRetainedIntronSequenceMap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 10,
  fig.height = 5,
  out.width = "100%"
)
```

## Overview

`createRetainedIntronSequenceMap()` measures per-position sequence motif
frequency at retained-intron junctions using the same two-bin window as
`createRetainedIntronSplicingMap()`. It accepts the same parameters as
`createSequenceMap()` with `RIMATS` in place of `SEMATS`.

```{r load}
library(RNAPeaks)
```

---

## Basic usage

```{r basic, eval = FALSE}
library(BSgenome.Hsapiens.UCSC.hg38)

createRetainedIntronSequenceMap(
    RIMATS   = sample_se.mats,
    sequence = "YCAY"
)
```

---

## Multiple motifs

```{r multi, eval = FALSE}
plots <- createRetainedIntronSequenceMap(
    RIMATS   = sample_se.mats,
    sequence = c("YCAY", "YGCY")
)
plots[["YCAY"]]
```

---

## Using a custom genome

```{r custom-genome, eval = FALSE}
library(BSgenome.Mmusculus.UCSC.mm10)

createRetainedIntronSequenceMap(
    RIMATS   = sample_se.mats,
    sequence = "YCAY",
    genome   = BSgenome.Mmusculus.UCSC.mm10
)
```

---

## IUPAC codes and returning data

The full IUPAC code table and `return_data` behavior are identical to
`createSequenceMap()`. See the
[createSequenceMap article](sequence-map.html) for details.

