---
title: "Getting Started with RNAPeaks"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Getting Started with RNAPeaks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 8,
  fig.height = 5,
  out.width = "100%"
)
```

## Overview

RNAPeaks is an R package for producing publication-quality visualizations
of RNA-binding protein (RBP) peaks overlaid on annotated gene structures.
Starting from a BED file of protein binding sites, RNAPeaks draws each peak
as a rectangle above the corresponding gene model — showing exons, UTRs, and
introns at accurate genomic scale.

The package provides four analysis modes:

| Mode | Function | Input |
|------|----------|-------|
| Single-gene peak plot | `PlotGene()` | BED + gene ID |
| Genomic region peak plot | `PlotRegion()` | BED + coordinates |
| Splicing map | `createSplicingMap()` | BED + SE.MATS |
| Sequence motif map | `createSequenceMap()` | SE.MATS + motif |

---

## Installation

```{r install, eval = FALSE}
# Bioconductor dependencies
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(c(
    "GenomicRanges", "IRanges", "S4Vectors", "BiocGenerics",
    "GenomeInfoDb", "AnnotationHub", "Biostrings", "BSgenome"
))

# Install RNAPeaks from GitHub
devtools::install_github("Krushna-B/RNAPeaks")
```

---

## Loading the package

```{r load}
library(RNAPeaks)
```

---

## Loading GTF annotation

Gene structure diagrams are built from an Ensembl GTF annotation. Load it
once per session with `LoadGTF()` and optionally cache it to disk to avoid
re-downloading on subsequent sessions.

```{r load-gtf, eval = FALSE}
# First call downloads from AnnotationHub (~1–2 min)
gtf <- LoadGTF(species = "Human")

# Cache locally for instant loading next time
saveRDS(gtf, "human_gtf.rds")
```

```{r load-gtf-bundled}
# The package ships with a pre-loaded GTF — no download required
data(gtf_human)
gtf <- gtf_human
```

Supported species: `"Human"` (Ensembl GRCh38, AH110867) and
`"Mouse"` (Ensembl GRCm38, AH47076).

---

## Preparing a BED file

Use `checkBed()` to validate and normalize any BED-format data frame before
passing it to a plotting function. Columns are mapped **by position**:

| Position | Name | Default |
|----------|------|---------|
| 1 | `chr` | required |
| 2 | `start` | required |
| 3 | `end` | required |
| 4 | `tag` | `"peak"` |
| 5 | `score` | `0` |
| 6 | `strand` | `"+"` |

```{r check-bed, eval = FALSE}
my_bed <- read.table("my_peaks.bed", header = FALSE, sep = "\t")
my_bed <- checkBed(my_bed)
```

The package includes a ready-to-use sample dataset:

```{r sample-bed}
head(sample_bed)
```

---

## Gene-level visualization

`PlotGene()` plots all RBP peaks that overlap a single gene of interest.

```{r plot-gene, eval = FALSE}
result <- PlotGene(
    bed    = sample_bed,
    geneID = "GAPDH",
    gtf    = gtf
)

result$plot   # display the plot
result$csv    # filtered peaks used in the figure
```

**Key parameters:**

| Parameter | Default | Description |
|-----------|---------|-------------|
| `geneID` | — | Gene symbol or Ensembl ID (`"ENSG..."`) |
| `TxID` | `NA` | Specific transcript isoform; defaults to longest |
| `species` | `"Human"` | `"Human"` or `"Mouse"` |
| `order_by` | `"Count"` | Peak track ordering: `"Count"`, `"Target"`, or `"Region"` |
| `merge` | `0` | Merge peaks within this many bp of each other |
| `five_to_three` | `FALSE` | Orient plot 5′ → 3′ regardless of strand |

---

## Region-level visualization

`PlotRegion()` displays all genes and their overlapping peaks within a
specified genomic window.

```{r plot-region, eval = FALSE}
result <- PlotRegion(
    bed    = sample_bed,
    gtf    = gtf,
    Chr    = "12",
    Start  = 56000000,
    End    = 56050000,
    Strand = "+"
)
result$plot
```

---

## Splicing map analysis

`createSplicingMap()` quantifies RBP binding frequency at every position
across a four-region window spanning exon-intron junctions of skipped-exon
(SE) events from rMATS output.

```{r splicing-map, eval = FALSE}
createSplicingMap(
    bed_file  = sample_bed,
    SEMATS    = sample_se.mats
)
```

The function classifies events into three groups based on `IncLevelDifference`
and `PValue` / `FDR`:

- **Retained** — significantly more inclusion in condition 1
- **Excluded** — significantly more skipping in condition 1
- **Control** — non-significant, stable inclusion

A bootstrap procedure samples a matched number of control events across
`control_iterations` replicates to build a null distribution. Positions where
Retained or Excluded frequencies exceed the control z-score threshold are
highlighted as significant regions.

```{r splicing-map-advanced, eval = FALSE}
createSplicingMap(
    bed_file           = sample_bed,
    SEMATS             = sample_se.mats,
    cores              = 4,           # parallel processing
    control_iterations = 20,          # bootstrap replicates
    show_significance  = TRUE,        # overlay significance bars
    WidthIntoExon      = 50,          # bp into flanking exons
    WidthIntoIntron    = 300          # bp into flanking introns
)
```

---

## Sequence motif analysis

`createSequenceMap()` works identically to `createSplicingMap()` but instead
of protein binding, it measures the per-position frequency of a target
sequence motif. Full IUPAC ambiguity codes are supported.

```{r seq-map, eval = FALSE}
library(BSgenome.Hsapiens.UCSC.hg38)

# YCAY motif (Y = C or T; classic NOVA binding site)
createSequenceMap(
    SEMATS   = sample_se.mats,
    sequence = "YCAY"
)
```

Common IUPAC codes:

| Code | Bases |
|------|-------|
| `R` | A or G |
| `Y` | C or T |
| `S` | G or C |
| `W` | A or T |
| `N` | any base |

---

## Session info

```{r session-info}
sessionInfo()
```
