---
title: "createSequenceMap"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{createSequenceMap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 10,
  fig.height = 5,
  out.width = "100%"
)
```

## Overview

`createSequenceMap()` measures per-position sequence motif frequency across
the same four-region window used by `createSplicingMap()`, applied to
skipped-exon (SE) events. Full IUPAC ambiguity codes are supported, enabling
analysis of degenerate RBP binding motifs.

The analysis logic, event classification, bootstrap procedure, and significance
testing are identical to `createSplicingMap()`. See the
[createSplicingMap article](splicing-maps.html) for details on those
parameters.

```{r load}
library(RNAPeaks)
```

---

## Basic usage

```{r basic, eval = FALSE}
library(BSgenome.Hsapiens.UCSC.hg38)

createSequenceMap(
    SEMATS   = sample_se.mats,
    sequence = "YCAY"    # NOVA binding motif; Y = C or T
)
```

---

## IUPAC ambiguity codes

Any standard IUPAC code is supported:

| Code | Matches |
|------|---------|
| `R` | A, G |
| `Y` | C, T |
| `S` | G, C |
| `W` | A, T |
| `K` | G, T |
| `M` | A, C |
| `B` | C, G, T |
| `D` | A, G, T |
| `H` | A, C, T |
| `V` | A, C, G |
| `N` | A, C, G, T |

```{r iupac, eval = FALSE}
# YGCY motif (MBNL binding site)
createSequenceMap(SEMATS = sample_se.mats, sequence = "YGCY")

# Poly-C run
createSequenceMap(SEMATS = sample_se.mats, sequence = "CCCC")
```

---

## Multiple motifs

Pass a character vector to `sequence` to generate one plot per motif:

```{r multi, eval = FALSE}
plots <- createSequenceMap(
    SEMATS   = sample_se.mats,
    sequence = c("YCAY", "YGCY", "CCCC")
)
plots[["YCAY"]]
```

---

## Using a custom genome

By default `createSequenceMap()` uses `BSgenome.Hsapiens.UCSC.hg38`. Supply
any installed BSgenome object to use a different assembly:

```{r custom-genome, eval = FALSE}
library(BSgenome.Mmusculus.UCSC.mm10)

createSequenceMap(
    SEMATS   = sample_se.mats,
    sequence = "YCAY",
    genome   = BSgenome.Mmusculus.UCSC.mm10
)
```

---

## Returning data

```{r return-data, eval = FALSE}
freq_df <- createSequenceMap(
    SEMATS      = sample_se.mats,
    sequence    = "YCAY",
    return_data = TRUE
)
head(freq_df)
```

