---
title: "Splicing Maps and Sequence Motif Analysis"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Splicing Maps and Sequence Motif Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 10,
  fig.height = 5,
  out.width = "100%"
)
```

## Overview

`createSplicingMap()` and `createSequenceMap()` quantify signal — protein
binding or sequence motif frequency — at every position across a four-region
window that spans the exon-intron junctions of skipped-exon (SE) alternative
splicing events.

Both functions accept SE.MATS output from [rMATS](https://rnaseq-mats.sourceforge.net/)
and classify events into three groups for comparison:

- **Retained** — exon inclusion significantly increases in condition 1
- **Excluded** — exon skipping significantly increases in condition 1
- **Control** — events with stable inclusion levels across conditions

```{r load}
library(RNAPeaks)
```

---

## The four-region window

Each splicing event is decomposed into four equal-length windows of
`(WidthIntoExon + WidthIntoIntron)` positions:

```
 Upstream exon        Intron 1       Skipped exon      Intron 2     Downstream exon
|──────|──────────────|──────|────────────────|──────|──────────────|──────|
  UE    ←Region 1→    UI3    ←──Region 2──→   EX5    ←─Region 3─→   DI3    ←Region 4→
       (UE-UI5)              (UI3-EX3)               (EX5-DI5)             (DI3-DE)
```

The x-axis of the output plot shows positions within this linearized window,
with vertical dashed lines at region boundaries and a schematic of the
skipped exon drawn beneath the x-axis.

---

## Splicing map

### Basic usage

```{r splicing-basic, eval = FALSE}
createSplicingMap(
    bed_file = sample_bed,
    SEMATS   = sample_se.mats
)
```

### Event classification thresholds

Control which events fall into each group:

```{r splicing-thresholds, eval = FALSE}
createSplicingMap(
    bed_file                        = sample_bed,
    SEMATS                          = sample_se.mats,
    p_valueRetainedAndExclusion     = 0.05,   # significance threshold for R/E
    p_valueControls                 = 0.95,   # min p-value for controls
    retained_IncLevelDifference     = 0.1,    # min ΔPSI for Retained
    exclusion_IncLevelDifference    = -0.1,   # max ΔPSI for Excluded
    Min_Count                       = 50,     # min junction read count
    use_fdr                         = TRUE    # use FDR instead of raw p-value
)
```

### Window size

Adjust how far into exons and introns the analysis extends:

```{r splicing-window, eval = FALSE}
createSplicingMap(
    bed_file        = sample_bed,
    SEMATS          = sample_se.mats,
    WidthIntoExon   = 75,    # 75 bp into each flanking exon
    WidthIntoIntron = 400    # 400 bp into each flanking intron
)
```

### Smoothing

A moving average is applied to reduce positional noise. Adjust the window
size or disable smoothing entirely:

```{r smoothing, eval = FALSE}
# Wider smoothing window
createSplicingMap(bed_file = sample_bed, SEMATS = sample_se.mats, moving_average = 100)

# No smoothing
createSplicingMap(bed_file = sample_bed, SEMATS = sample_se.mats, moving_average = NULL)
```

### Parallel processing

For large datasets, use multiple cores to speed up the binding frequency
calculation:

```{r parallel, eval = FALSE}
createSplicingMap(
    bed_file = sample_bed,
    SEMATS   = sample_se.mats,
    cores    = 4
)
```

### Bootstrap control and significance testing

The control distribution is built by bootstrapping: at each of
`control_iterations` replicates, a random sample of `(n_retained + n_excluded)
× control_multiplier` control events is drawn and their binding frequency
computed. The mean ± SD across replicates forms the control ribbon.

Significance is then assessed at each position using a z-score:

```
z = (observed_frequency − control_mean) / control_SD
```

Contiguous runs of ≥ `min_consecutive` significant positions are reported
as enriched regions and drawn as colored bars above the plot.

```{r significance, eval = FALSE}
createSplicingMap(
    bed_file           = sample_bed,
    SEMATS             = sample_se.mats,
    control_multiplier = 2.0,   # sample_size = (R + E) * 2
    control_iterations = 30,    # more iterations = tighter control band
    z_threshold        = 1.96,  # p < 0.05 (used when use_fdr = FALSE)
    min_consecutive    = 10,    # minimum run length to call a region
    one_sided          = TRUE,  # test enrichment only (not depletion)
    show_significance  = TRUE
)
```

### Validating bootstrap normality

The z-score test assumes that bootstrap replicates are approximately normally
distributed. Verify this with `return_diagnostics = TRUE`:

```{r diagnostics, eval = FALSE}
diag <- createSplicingMap(
    bed_file           = sample_bed,
    SEMATS             = sample_se.mats,
    return_diagnostics = TRUE
)

# Fraction of positions where bootstrap distribution passes Shapiro-Wilk
mean(
  apply(diag$bootstrap_matrix, 1, function(x) shapiro.test(x)$p.value) > 0.05
)
```

### Returning data

```{r return-data, eval = FALSE}
freq_df <- createSplicingMap(
    bed_file    = sample_bed,
    SEMATS      = sample_se.mats,
    return_data = TRUE
)
head(freq_df)
```

---

## Sequence motif analysis

`createSequenceMap()` accepts the same SE.MATS input and parameters as
`createSplicingMap()`, but instead of protein binding sites it measures the
per-position frequency of a target nucleotide motif in the genomic sequence.

### Basic usage

```{r seq-basic, eval = FALSE}
library(BSgenome.Hsapiens.UCSC.hg38)

createSequenceMap(
    SEMATS   = sample_se.mats,
    sequence = "YCAY"    # NOVA binding motif; Y = C or T
)
```

### IUPAC ambiguity codes

Any standard IUPAC ambiguity code is supported:

| Code | Matches |
|------|---------|
| `R` | A, G |
| `Y` | C, T |
| `S` | G, C |
| `W` | A, T |
| `K` | G, T |
| `M` | A, C |
| `B` | C, G, T |
| `D` | A, G, T |
| `H` | A, C, T |
| `V` | A, C, G |
| `N` | A, C, G, T |

```{r iupac, eval = FALSE}
# YGCY motif (MBNL binding site)
createSequenceMap(SEMATS = sample_se.mats, sequence = "YGCY")

# Poly-C run
createSequenceMap(SEMATS = sample_se.mats, sequence = "CCCC")
```

### Supplying a custom genome

By default `createSequenceMap()` uses `BSgenome.Hsapiens.UCSC.hg38`. Supply
any BSgenome object to use a different assembly:

```{r custom-genome, eval = FALSE}
library(BSgenome.Mmusculus.UCSC.mm10)

createSequenceMap(
    SEMATS   = sample_se.mats,
    sequence = "YCAY",
    genome   = BSgenome.Mmusculus.UCSC.mm10
)
```

---

## Appearance options

Both functions share a common set of visual parameters:

| Parameter | Default | Description |
|-----------|---------|-------------|
| `retained_col` | `"blue"` | Retained group line color |
| `excluded_col` | `"red"` | Excluded group line color |
| `control_col` | `"black"` | Control group line color |
| `line_width` | `0.8` | Line width |
| `ribbon_alpha` | `0.3` | Opacity of control SD ribbon |
| `boundary_col` | `"gray70"` | Region boundary line color |
| `exon_col` | `"navy"` | Skipped exon fill in the schematic |
| `legend_position` | `"bottom"` | Legend position |
| `ylab` | `"Frequency"` | Y-axis label |
| `title` | `""` | Plot title |

---

## Session info

```{r session-info}
sessionInfo()
```
