---
title: "createSplicingMap"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{createSplicingMap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 10,
  fig.height = 5,
  out.width = "100%"
)
```

## Overview

`createSplicingMap()` quantifies RBP binding frequency at every position
across a four-region window spanning the exon-intron junctions of skipped-exon
(SE) alternative splicing events from rMATS output.

```{r load}
library(RNAPeaks)
```

---
## Basic usage

```{r basic, eval = FALSE}
createSplicingMap(
    bed_file = sample_bed,
    SEMATS   = sample_se.mats
)
```

---

## Event classification

Events are classified into three groups based on `IncLevelDifference`,
`PValue`, and `FDR`:

- **Retained** — `IncLevelDifference < -0.1`: exon inclusion significantly higher in condition 1
- **Excluded** — `IncLevelDifference > 0.1`: exon skipping significantly higher in condition 1
- **Control** — non-significant, stable inclusion (PValue > 0.95)

Control which events fall into each group:

```{r thresholds, eval = FALSE}
createSplicingMap(
    bed_file                     = sample_bed,
    SEMATS                       = sample_se.mats,
    p_valueRetainedAndExclusion  = 0.05,
    p_valueControls              = 0.95,
    retained_IncLevelDifference  = -0.1,
    exclusion_IncLevelDifference = 0.1,
    Min_Count                    = 50
)
```

---

## Window size

Adjust how far into exons and introns the analysis extends:

```{r window, eval = FALSE}
createSplicingMap(
    bed_file        = sample_bed,
    SEMATS          = sample_se.mats,
    WidthIntoExon   = 75,
    WidthIntoIntron = 400
)
```

---

## Smoothing

A per-bin moving average is applied to reduce positional noise. Adjust the
window size or disable smoothing:

```{r smoothing, eval = FALSE}
# Wider smoothing window
createSplicingMap(bed_file = sample_bed, SEMATS = sample_se.mats, moving_average = 100)

# No smoothing
createSplicingMap(bed_file = sample_bed, SEMATS = sample_se.mats, moving_average = NULL)
```

---

## Bootstrap control and significance testing

The control distribution is built by bootstrapping: at each of
`control_iterations` replicates, a random sample of
`(n_retained + n_excluded) × control_multiplier` control events is drawn and
their binding frequency computed. The mean ± SD across replicates forms the
control ribbon on the plot.

Significance is assessed at each position using a z-score:

```
z = (observed_frequency − control_mean) / control_SD
```

Contiguous runs of ≥ `min_consecutive` significant positions are reported as
enriched regions and drawn as colored bars above the plot.

```{r significance, eval = FALSE}
createSplicingMap(
    bed_file           = sample_bed,
    SEMATS             = sample_se.mats,
    control_multiplier = 2.0,
    control_iterations = 30,
    use_fdr            = TRUE,
    fdr_threshold      = 0.05,
    min_consecutive    = 10,
    one_sided          = TRUE,
    show_significance  = TRUE
)
```

---

## Returning data

```{r return-data, eval = FALSE}
freq_df <- createSplicingMap(
    bed_file    = sample_bed,
    SEMATS      = sample_se.mats,
    return_data = TRUE
)
head(freq_df)
```

