---
title: "Styling Guide"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Styling Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 10,
  fig.height = 6,
  out.width = "100%"
)
```

## Overview

This article documents all visual styling parameters across the RNAPeaks
function suite. Parameters are grouped by function family.

```{r load}
library(RNAPeaks)
data(gtf_human)
```

---

## PlotGene and PlotRegion

`PlotGene()` and `PlotRegion()` share a common set of styling arguments.

### Gene structure

| Parameter | Default | Description |
|-----------|---------|-------------|
| `exon_fill` | `"navy"` | Exon / CDS fill color |
| `utr_fill` | `"lightgray"` | UTR fill color |
| `utr_col` | `"darkgray"` | UTR border color |
| `intron_color` | `"gray60"` | Intron line color |
| `intron_linewidth` | `0.9` | Intron line width |

### Peaks

| Parameter | Default | Description |
|-----------|---------|-------------|
| `peak_col` | `"purple"` | Peak rectangle fill color |
| `peak_alpha` | `0.95` | Peak opacity (0–1) |
| `peak_border_color` | `NA` | Peak border color (`NA` = none) |
| `peaks_width` | `0.3` | Vertical height of each peak row |

### Labels

| Parameter | Default | Description |
|-----------|---------|-------------|
| `label_size` | `5` | Protein track label font size |
| `label_color` | `"black"` | Protein track label color |
| `title_size` | `25` | Plot title font size |
| `bam_label_size` | `8` | BAM coverage track label size |

### Layout

| Parameter | Default | Description |
|-----------|---------|-------------|
| `axis_pad_bp` | `500` | Padding (bp) added to each side of the x-axis |
| `axis_breaks_n` | `5` | Number of x-axis tick marks |
| `max_proteins` | `40` | Maximum number of protein tracks shown |
| `x_lims` | `NULL` | Custom x-axis limits as `c(min, max)` |

### Background bands

| Parameter | Default | Description |
|-----------|---------|-------------|
| `band_even_fill` | `"#F7F8FA"` | Even-row band fill |
| `band_odd_fill` | `"#FFFFFF"` | Odd-row band fill |
| `band_sep_color` | `"#E5E7EB"` | Band separator line color |

### Example

```{r vis-style, eval = FALSE}
PlotGene(
    bed             = sample_bed,
    geneID          = "GAPDH",
    gtf             = gtf_human,
    exon_fill       = "steelblue",
    utr_fill        = "lightblue",
    peak_col        = "firebrick",
    peak_alpha      = 0.8,
    label_size      = 4,
    band_even_fill  = "#F0F4FF",
    band_odd_fill   = "#FFFFFF"
)
```

---

## createSplicingMap, createRetainedIntronSplicingMap, createSequenceMap, createRetainedIntronSequenceMap

All four analysis functions share the same visual parameters.

### Line colors

| Parameter | Default | Description |
|-----------|---------|-------------|
| `retained_col` | `"blue"` | Retained group line color |
| `excluded_col` | `"red"` | Excluded group line color |
| `control_col` | `"black"` | Control group line color |

### Lines and ribbon

| Parameter | Default | Description |
|-----------|---------|-------------|
| `line_width` | `0.8` | Frequency line width |
| `line_alpha` | `0.7` | Frequency line opacity |
| `ribbon_alpha` | `0.3` | Opacity of the control SD ribbon |

### Schematic

| Parameter | Default | Description |
|-----------|---------|-------------|
| `boundary_col` | `"gray70"` | Region boundary dashed line color |
| `exon_col` | `"navy"` | Skipped / flanking exon fill in the schematic |

### Title and axes

| Parameter | Default | Description |
|-----------|---------|-------------|
| `title` | `""` | Plot title |
| `title_size` | `20` | Plot title font size |
| `title_color` | `"black"` | Plot title color |
| `axis_text_size` | `11` | Y-axis tick label font size |
| `ylab` | `"Frequency"` | Y-axis label |

### Legend

| Parameter | Default | Description |
|-----------|---------|-------------|
| `legend_position` | `"bottom"` | Legend position: `"bottom"`, `"top"`, `"left"`, `"right"`, or `"none"` |

### Example

```{r map-style, eval = FALSE}
createSplicingMap(
    bed_file        = sample_bed,
    SEMATS          = sample_se.mats,
    retained_col    = "#1f77b4",
    excluded_col    = "#d62728",
    control_col     = "#7f7f7f",
    ribbon_alpha    = 0.2,
    exon_col        = "steelblue",
    boundary_col    = "gray50",
    title           = "RBP Binding Around SE Junctions",
    title_size      = 18,
    ylab            = "Binding Frequency",
    legend_position = "right"
)
```

